---
title: "Rank IPL Players"
author: "Tinniam V Ganesh"
date: "April 24, 2016"
output: html_document
---

Here is a short post which ranks IPL batsmen and bowlers. These are based on match data
from [Cricsheet](http://www.cricsheet.org). Ranking batsmen and bowlers in IPL is more challenging as the players can belong to different teams in different years. Hence I create a combined data frame of the batsmen and bowlers regardless of their IPL teams and calculate  a) average runs and average strike rate for batsmen and c) average wickets and d) average economy rate for bowlers.

I will be doing this ranking for T20 and ODI batting and bowling performances shortly.

This post has also been published in RPubs [RankIPLPlayers](http://rpubs.com/tvganesh/RankIPLPlayers). You can download this as a pdf file at [RankIPLPlayers.pdf](https://drive.google.com/file/d/0Bxz66bp5PxPBT3ZFaGxvT1NnTUU/view?usp=sharing)

The results are slightly surprising

```{r message=FALSE,warning=FALSE}
rm(list=ls())
library(yorkr)
library(dplyr)
setwd("C:/software/cricket-package/cricsheet/cleanup/IPL/rank")
source("rankIPLBatsmen.R")
source("rankIPLBowlers.R")
```

## Rank IPL batsmen
Chris Gayle, MEK Hussey and Shane Watson are top 3 IPL batsmen. Gayle towers over the others in mean runs and mean strike rate. Surprisingly Ajinkya Rahane is the top Indian T20 batsman, if we leave out Sachin Tendulkar (who tops India yet again!). The other top IPL T20 batsmen are Raina, Gambhir, Rohit Sharma in that order. Virat Kohli comes a distant 14th.
```{r }
iplBatsmanRank <- rankIPLBatsmen()
as.data.frame(iplBatsmanRank[1:30,])
```

## Rank IPL bowlers
The top 3 IPL T20 bowlers are Sunil Narine, SL Malinga and Anil Kumble. Munaf patel, Ashish Nehra and RP Singh are the other top IPL T20 bowlers.RA Jadeja is 24th and R Ashwin is 31st.

Don't get hung up on the decimals in the average wickets for the bowlers. All it implies is that if 2 bowlers have average wickets of 1.0 and 1.5, it implies that in 2 matches the 1st bowler will take 2 wickets and the 2nd bowler will take 3 wickets. 
```{r}
setwd("C:/software/cricket-package/cricsheet/cleanup/IPL/rank")
iplBowlersRank <- rankIPLBowlers()
as.data.frame(iplBowlersRank[1:31,])
```


Conclusion: The results are somewhat surprising. The ranking was based on data from
[Cricsheet](http://www.cricsheet.org). The data in this site are available from 2008-2015.
I hope to do this ranking for T20 and ODIs shortly

Watch this space!

1. [Introducing cricket package yorkr-Part1:Beaten by sheer pace!](https://gigadom.wordpress.com/2016/04/02/introducing-cricket-package-yorkr-part-1-beaten-by-sheer-pace/). 
2. [yorkr pads up for the Twenty20s: Part 1- Analyzing team"s match performance](https://gigadom.wordpress.com/2016/04/16/yorkr-pads-up-for-the-twenty20s-part-1-analyzing-teams-match-performance/).
3. [yorkr crashes the IPL party !Part 1](https://gigadom.wordpress.com/2016/04/19/yorkr-crashes-the-ipl-party-part-1/)
4. [Introducing cricketr! : An R package to analyze performances of cricketers](https://gigadom.wordpress.com/2015/07/04/introducing-cricketr-a-r-package-to-analyze-performances-of-cricketers/)
4. [Cricket analytics with cricketr in paperback and Kindle versions](https://gigadom.wordpress.com/2016/02/05/cricket-analytics-with-cricketr-in-paperback-and-kindle-versions/)